---
title: "Anomalous Cases Detection Report"
author: ""
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
params:
  mlAlgorithm : ""
  geneticAlgorithm : NULL
  lassoPredictors : NULL
  baselinePrecision : NULL
  baselinePredictors : NULL
  originalDiagnosis : NULL
  clinicData : NULL
  categoricActivePredictors : NULL,
  numericActivePredictors : NULL,
  selectedData : NULL
  pcaAnalysis : NULL
  pcaAnalysisSelected : NULL
  pcaInfo: NULL
  numericTable : NULL
  totalTable : NULL
  classVariable: NULL
  idColumn : NULL
  predictorsToSelect : NULL
  numTrees : NULL
  mtry : NULL
  splitRule : NULL 
  sampleFraction : NULL
  maxDepth : NULL
  minNodeSize : NULL
  nIterations : NULL 
  nStopIter : NULL
  populationSize : NULL 
  diagnosticChangeProbability : NULL 
  crossoverOperator : NULL
  crossoverProbability : NULL 
  selectionOperator : NULL
  mutationOperator : NULL
  mutationProbability : NULL 
  nCores : NULL
  seed : NULL
  numModelExecutions : NULL
  bestAfterDetectionCM : NULL
  bestBaselineModelCM : NULL
  predictorsImp : NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(plotly)
library(lazyeval)
library(ggtext)
```

```{r, include=FALSE}
modelName <- ""

if (mlAlgorithm == "Lasso"){
  modelName <- "Lasso"
  executeLasso <- TRUE
  executeRF <- FALSE
} else {
  modelName <- "Random Forest"
  executeLasso <- FALSE
  executeRF <- TRUE
}

numExecutions <- length(params$geneticAlgorithm@summary[,1])

maxPrecision <- round(params$geneticAlgorithm@fitnessValue * 100, 2)

baselinePrecision <- round(params$baselinePrecision * 100, 2)

precisionLine <- round(params$baselinePrecision * 100, 0)

predictorsLine <- params$baselinePredictors

lassoText <- paste(paste("Also, the mean number of selected predictors has gone from", predictorsLine, "to", tail(params$lassoPredictors, n=1),  sep = " "), ".", sep = "")

bestAfterDetectionCM <- params$bestAfterDetectionCM
bestBaselineModelCM <- params$bestBaselineModelCM
```

```{r, include = FALSE}
classVariable <- params$classVariable
idColumn <- params$idColumn

nCores <- params$nCores
seed <- params$seed

populationSize <- params$populationSize
diagnosticChangeProbability <- params$diagnosticChangeProbability
nIterations <- params$nIterations
nStopIter <- params$nStopIter
selectionOperator <- params$selectionOperator
crossoverOperator <- params$crossoverOperator
crossoverProbability <- params$crossoverProbability
mutationOperator <- params$mutationOperator
mutationProbability <- params$mutationProbability

numModelExecutions <- params$numModelExecutions


if(is.null(params$numericActivePredictors)){
  
  numericVariables <- rownames(params$numericTable)

} else {
  
  numericVariables <- params$numericActivePredictors
}
 

if(is.null(params$categoricActivePredictors)){
  
  clinicNames <- names(params$clinicData)
  
  categoricVariables <- clinicNames[!(clinicNames %in% numericVariables)]
  
} else {
  
  categoricVariables <- params$categoricActivePredictors
}
```

# 1. Parameters and hyperparameters used for anomalous case detection

The class variable used in the anomalous sample detection process was **`r classVariable`**. The variable used as identifier of each patient has been **`r idColumn`**.

This process has been performed by applying a parallelization with **`r nCores`** cores. In addition, the seed used in the process has been **`r seed`**.

### 1.1 Parameters chosen for the genetic algorithm

The choice of parameters for the execution of the genetic algorithm was as follows:

- Size of the initial population: **`r populationSize`**.
- Percentage of patients whose diagnosis is changed at the time of generating the initial population: **`r diagnosticChangeProbability`**.

- Maximum number of iterations to perform: **`r nIterations`**.
- Number of iterations to stop if the fitness value doesn't change: **`r nStopIter`**.

- Selection operator: **`r selectionOperator`**.

- Crossover operator: **`r crossoverOperator`**.
- Probability of crossover: **`r crossoverProbability`**.

- Mutation operator: **`r mutationOperator`**.
- Probability of mutation: **`r mutationProbability`**.


### 1.2 Model selection mechanism chosen
Regarding the model selection mechanism:

- The hyperparameter grid used contains a unique combination.
- The type of validation used was repetition, specifically with **`r numModelExecutions`** repetitions.
 
### 1.3 Algorithm chosen and the hyperparameters chosen for it
A `r modelName` model has been used to detect anomalous cases.

`r if(executeLasso){"No hyperparameter is chosen for this model, since the alpha is chosen by R."}`

`r if(executeRF){"The following hyperparameters have been chosen for this model:"}`

`r if(executeRF){paste("- Number of trees of the Random Forest model: ", "**", params$numTrees, "**.", sep = "")}`
`r if(executeRF){paste("- Number of predictors that are evaluated at each partition (node) of each tree: ", "**", params$mtry, "**.", sep = "")}`
`r if(executeRF){paste("- Rule used to select the predictor and the optimal value to separate a node into two branches during the tree construction: ", "**", params$splitRule, "**.", sep = "")}`
`r if(executeRF){paste("- Fraction of the training data that will be used to create each of the trees in the forest: ", "**", params$sampleFraction, "**.", sep = "")}`
`r if(executeRF){paste("- Maximum height of each tree in the forest: ", "**", params$maxDepth, "**.", sep = "")}`
`r if(executeRF){paste("- Minimum number of observations required in a node to be able to split it: ", "**", params$minNodeSize, "**.", sep = "")}`

`r if(executeRF){paste("In the RF model it is necessary to indicate the number of predictors to be selected from the most important predictors. In this case the number of predictors selected has been: **", params$predictorsToSelect, "**." , sep = "")}`

### 1.4 Categorization of clinical data

The active predictors of numerical type are: `r paste("**", paste(numericVariables, collapse = ", "), "**", sep = "")`

The active predictors of categorical type are: `r paste("**", paste(categoricVariables, collapse = ", "), "**", sep = "")`

# 2. Evolution of the `r modelName` model during the detection

In the anomalous sample detection process, `r numExecutions` iterations have been performed. During this process, the accuracy has gone from `r baselinePrecision` percent to `r maxPrecision` percent.
`r if(executeLasso){lassoText}`

```{r, include = FALSE}
algorithmPrecisions <- params$geneticAlgorithm@summary[,1]

precisionDF <- data.frame(generation = 1:length(algorithmPrecisions), precision = algorithmPrecisions * 100, meanPrecision = cummean(algorithmPrecisions * 100))

hline <- function(y = 0, color = "black") {
  list(
    type = "line",
    x0 = 0,
    x1 = 1,
    xref = "paper",
    y0 = y,
    y1 = y,
    line = list(color = color)
  )
}
```

### 2.1 Evolution of the balanced accuracy obtained
```{r, include = FALSE}
precisionPlot <- plot_ly(precisionDF, x=~generation, y=~precision, hoveron = "points", hoverinfo = "text", text = paste("Maximum accuracy value obtained: ", round(precisionDF$precision, 2), "%<br>Generation: ", precisionDF$generation, sep = ""), type = 'scatter', mode = 'lines+markers', name ="Precision", color = I("cadetblue3"))
  
precisionPlot <- precisionPlot %>% add_trace(x=~generation, y = ~meanPrecision, name='Mean', mode = 'lines+markers', color = I("brown3"), hoverinfo = "text", text = paste("Average of the accuracies obtained: ", round(precisionDF$meanPrecision, 2), "%<br>Generation: ", precisionDF$generation, sep = "")) 
    
precisionPlot <- precisionPlot %>%  layout(title = "Evolution of the accuracy obtained during generations", xaxis = list(title = "Actual generation <br>"), yaxis = list(title = "Precision value <br> <br> <br>", ticksuffix = "%",  range = c(precisionLine - 5, min(100, (max(algorithmPrecisions) * 100) + 5)), xaxis = list(range = c(0, max(precisionDF$generation)))))

precisionPlot <- precisionPlot %>% layout(shapes = list(hline(precisionLine))) %>%
  add_annotations(x = length(algorithmPrecisions) - 3, y = precisionLine + 1, showarrow = F,
           text = paste("Baseline precision: ", precisionLine, "%", sep = ""))
```

```{r}
precisionPlot
```

`r if(executeLasso){"### 2.2 Evolution of the number of predictors selected"}`
```{r, include = FALSE, eval = executeLasso}
predictorsDF <- data.frame(generation = 1:length(params$lassoPredictors), predictors = params$lassoPredictors)

maxPredictors <- max(params$lassoPredictors)
minPredictors <- min(params$lassoPredictors)
  
predictorsPlot <- plot_ly(predictorsDF, x=~generation, y=~predictors, hoveron = "points", hoverinfo = "text", text = paste("Mean number of predictors selected: ", round(predictorsDF$predictors, 2), "<br>Generation: ", precisionDF$generation, sep = ""), type = 'scatter', mode = 'lines+markers', name ="Number of predictors", color = I("#FF7256"))
  
predictorsPlot <- predictorsPlot %>%  layout(title = "Evolution of the number of predictors selected during generations", xaxis = list(title = "Actual generation <br>"), yaxis = list(title = "Number of predictors <br> <br> <br>", range = c(min(c(predictorsLine, min(params$lassoPredictors))) - 5, max(c(predictorsLine, max(params$lassoPredictors))) + 5), xaxis = list(range = c(0, max(predictorsDF$generation)))))

predictorsPlot <- predictorsPlot %>% layout(shapes = list(hline(predictorsLine))) %>%
  add_annotations(x = length(algorithmPrecisions) - 3, y = predictorsLine + (((maxPredictors - minPredictors) / 10) / 4 ), showarrow = F,
           text = paste("Baseline selected predictors: ", predictorsLine, sep = ""))
```

```{r, include = executeLasso, eval = executeLasso}
predictorsPlot
```

`r if(executeLasso){"### 2.3 Confusion Matrix of the best model obtained before the detection"}`
`r if(!executeLasso){"### 2.2 Confusion Matrix of the best model obtained before the detection"}`
```{r}
bestBaselineModelCM
```

`r if(executeLasso){"### 2.4 Confusion Matrix of the best model obtained after the detection"}`
`r if(!executeLasso){"### 2.3 Confusion Matrix of the best model obtained after the detection"}`
```{r}
bestAfterDetectionCM
```

# 4. Diagnostics distribution before and after the detection
```{r, include = FALSE}
originalDF <- data.frame(
  Case = sum( params$originalDiagnosis == "Case"),
  Control = sum( params$originalDiagnosis == "Control")
)

changedDF <- data.frame(
  Case = sum(params$clinicData[[params$classVariable]] == "Case"),
  Control = sum(params$clinicData[[params$classVariable]] == "Control"),
  Case2Control = sum(params$clinicData[[params$classVariable]] == "Case2Control"),
  Control2Case = sum(params$clinicData[[params$classVariable]] == "Control2Case")
)
```

```{r, include = FALSE}
originalBars <- plot_ly(originalDF, x = "CASE", y = ~Case, type = 'bar', name = 'Num Case', marker = list(color = '#FF7256'), hovertext = ~Case, hovertemplate = "Patients: %{y}<extra></extra>")

originalBars <- originalBars %>% add_trace(x = "CONTROL", y = ~Control, name = 'Num Control', marker = list(color = '#CAFF70'), hovertext = ~Control, hovertemplate = "Patients: %{y}<extra></extra>")

originalBars <- originalBars %>% style(hoverinfo = 'none')

originalBars <- originalBars %>% layout(title = "Distribution of patient diagnoses", yaxis = list(title = "Number of patients (before detection) <br> <br>")) 
```

```{r, include = FALSE}
changedBars <- plot_ly(changedDF, x = "CASE", y = ~Case, type = 'bar', name = 'Num Case', marker = list(color = '#FF7256'), hovertext = ~Case, hovertemplate = "Patients: %{y}<extra></extra>")

changedBars <- changedBars %>% add_trace(x = "CONTROL", y = ~Control, name = 'Num Control', marker = list(color = '#CAFF70'), hovertext = ~Control, hovertemplate = "Patients: %{y}<extra></extra>")

changedBars <- changedBars %>% add_trace(x = "CASE2CONTROL", y = ~Case2Control, name = 'Num Case2Control', marker = list(color = '#FFB90F'), hovertext = ~Case2Control, hovertemplate = "Patients: %{y}<extra></extra>")

changedBars <- changedBars %>% add_trace(x = "CONTROL2CASE", y = ~Control2Case, name = 'Num Control2Case', marker = list(color = '#FF7F00'), hovertext = ~Control2Case, hovertemplate = "Patients: %{y}<extra></extra>")

changedBars <- changedBars %>% style(hoverinfo = 'none')

changedBars <- changedBars %>% layout(title = "Distribution of patient diagnoses", yaxis = list(title = "Number of patients (after detection) <br> <br>")) 
```

```{r, include = FALSE}
distributionDashboard <- subplot(originalBars, changedBars, nrows = 1, margin = 0.1, shareX = FALSE, shareY = FALSE, titleX = TRUE, titleY = TRUE)  

distributionDashboard <- distributionDashboard %>%  layout(title = "Distribution of patient diagnoses", showlegend = FALSE)
```

```{r}
distributionDashboard
```

# 5. PCA Analysis

```{r, include = FALSE}
pcaInfo <- params$pcaInfo
```

### 5.1 PCA Analysis over total predictors
```{r, include = FALSE}
classVariable <- params$pcaAnalysis[[params$classVariable]]

pcaAnalysis <- ggplot(params$pcaAnalysis, aes(PC1, PC2)) + 
    modelr::geom_ref_line(h = 0) +
    modelr::geom_ref_line(v = 0) +
    geom_jitter(aes(color = classVariable)) +
    xlab(paste("First Principal Component ->", pcaInfo["PC1"], "% variability explained", sep = "")) + 
    ylab(paste("Second Principal Component ->", pcaInfo["PC2"], "% variability explained", sep = "")) + 
    ggtitle(paste("PCA Analysis of Changed Diagnostics ", pcaInfo["PC1"] + pcaInfo["PC2"], "% variability explained", sep = ""))
```

```{r}
pcaAnalysis
```

### 5.2 PCA Analysis over the selected predictors by the model
```{r, include = FALSE}
pcaSelected <- ggplot(params$pcaAnalysisSelected, aes(PC1, PC2)) + 
    modelr::geom_ref_line(h = 0) +
    modelr::geom_ref_line(v = 0) +
    geom_jitter(aes(color = classVariable)) +
    xlab(paste("First Principal Component ->", pcaInfo["PC1Selected"], "% variability explained", sep = "")) + 
    ylab(paste("Second Principal Component ->", pcaInfo["PC2Selected"], "% variability explained", sep = "")) + 
    labs(title = "PCA Analysis of Changed Diagnostics in selected predictors<br>Another line")  +
    theme(plot.title.position = "plot",
        plot.title = element_markdown(),
        legend.position = "none")
```

```{r, eval=FALSE,include=FALSE}
ggtitle(paste("PCA Analysis of Changed Diagnostics in selected predictors<br>", pcaInfo["PC1Selected"] + pcaInfo["PC2Selected"], "% variability explained", sep = ""))
```


```{r}
pcaSelected
```

# 6. Importance of final selected predictors
```{r}
DT::datatable(params$predictorsImp, class = 'cell-border stripe', options = list(scrollX = TRUE))
```

# 7. Omics for the final selected predictors
```{r}
DT::datatable(params$selectedData, class = 'cell-border stripe', options = list(scrollX = TRUE))
```

# 8. Clinical data for individuals in the study
```{r}
DT::datatable(params$clinicData, class = 'cell-border stripe', options = list(scrollX = TRUE))
```

# 9. Ratio Analysis

```{r, include=FALSE}
numericTable <- params$numericTable

totalTable <- params$totalTable
```

```{r, include = FALSE}
totalTable$percentageCCCM <- totalTable$CCCM - 1
totalTable$percentageCCCM <- totalTable$percentageCCCM * 100

totalTable$percentageECEM <- totalTable$ECEM - 1
totalTable$percentageECEM <- totalTable$percentageECEM * 100

totalTable$percentageCE <- totalTable$CE - 1
totalTable$percentageCE <- totalTable$percentageCE * 100

totalTable$logCCCM <- ifelse(totalTable$CCCM == 0, 1, totalTable$CCCM)

totalTable$logCCCM <- log(totalTable$logCCCM, base = 2)

totalTable$logECEM <- ifelse(totalTable$ECEM == 0, 1, totalTable$ECEM)

totalTable$logECEM <- log(totalTable$logECEM, base = 2)

totalTable$logCE <- ifelse(totalTable$CE == 0, 1, totalTable$CE)

totalTable$logCE <- log(totalTable$logCE, base = 2)
```

```{r, include = FALSE}
maxCCCM <- max(totalTable$logCCCM)
minCCCM <- min(totalTable$logCCCM)

maxECEM <- max(totalTable$logECEM)
minECEM <- min(totalTable$logECEM)

maxCE <- max(totalTable$logCE)
minCE <- min(totalTable$logCE)

if(abs(maxCCCM) + abs(minCCCM) > 3){
  sepCCCM <- 0.5
} else if(abs(maxCCCM) + abs(minCCCM) > 1) {
  sepCCCM <- 0.25
} else {
  sepCCCM <- 0.1
}

if(abs(maxECEM) + abs(minECEM) > 3){
  sepECEM <- 0.5
} else if(abs(maxECEM) + abs(minECEM) > 1) {
  sepECEM <- 0.25
} else {
  sepECEM <- 0.1
}


if(abs(maxCE) + abs(minCE) > 3){
  sepCE <- 0.5
} else if(abs(maxCE) + abs(minCE) > 1) {
  sepCE <- 0.25
} else {
  sepCE <- 0.1
}
```


```{r, include=FALSE}
totalTable$ColorsCCCM <- NA
totalTable$ColorsECEM <- NA
totalTable$ColorsCE <- NA

totalTable$textCCCM <- NA
totalTable$textECEM <- NA
totalTable$textCE <- NA

for(indexRatio in 1:nrow(totalTable)){

  if(totalTable$CCCM[indexRatio] == 0){

    totalTable$ColorsCCCM[indexRatio] <- "black"

  } else {

    if(abs(totalTable$logCCCM[indexRatio]) < 0.5){
      totalTable$ColorsCCCM[indexRatio] <- "gold"
    } else if (abs(totalTable$logCCCM[indexRatio]) > 0.5 && abs(totalTable$logCCCM[indexRatio]) < 1.5){
      totalTable$ColorsCCCM[indexRatio] <- "orange"
    } else {
      totalTable$ColorsCCCM[indexRatio] <- "red"
    }

  }


  text <- ""
  percentageText <- ""
  
  if(rownames(totalTable)[indexRatio] %in% numericVariables){

    text <- strsplit(rownames(totalTable)[indexRatio], " vs ")[[1]][1]
    text <- paste(text, "(Numeric)", "<br>", sep = " ")
    
    text <- paste(text, "Ratio: ", round(totalTable$CCCM[indexRatio], digits = 3), " -> (", round(numericTable[rownames(totalTable)[indexRatio], "MeanCC"], digits = 3), "/", round(numericTable[rownames(totalTable)[indexRatio], "MeanCM"], digits = 3), ")",  sep = "")
    
    text <- paste(text, "Ratio (logaritmic scale):", sep = "<br>")
    text <- paste(text, round(totalTable$logCCCM[indexRatio], digits = 3),  sep = " ")
    
    text <- paste(text, "Number of Control -> Case patients:", sep = "<br>")
    text <- paste(text, numericTable[rownames(totalTable)[indexRatio], "NumPatientsCC"], sep = " ")
    
    text <- paste(text, "Mean value for the Control -> Case patients:", sep = "<br>")
    text <- paste(text, round(numericTable[rownames(totalTable)[indexRatio], "MeanCC"], digits = 3), sep = " ")

    text <- paste(text, "Number of Control -> Control patients:", sep = "<br>")
    text <- paste(text, numericTable[rownames(totalTable)[indexRatio], "NumPatientsCM"], sep = " ")

    text <- paste(text, "Mean value for the Control -> Control patients:", sep = "<br>")
    text <- paste(text, round(numericTable[rownames(totalTable)[indexRatio], "MeanCM"], digits = 3), sep = " ")
    
    percentageText <- paste("The variable **", rownames(totalTable)[indexRatio], "** is **numeric**.", sep = "")
      
    percentageText <- paste(percentageText, "\n", sep ="")
    
    percentageText <- paste(percentageText, "A **", round(totalTable$CCCM[indexRatio], digits = 3), " ratio** is obtained.", sep = "")
    
    
    if(totalTable$CCCM[indexRatio] != 0){
      
      percentageText <- paste(percentageText, "This ratio means:", sep = " ")
    
      percentageText <- paste(percentageText, "\n", sep ="")
      
      if (totalTable$percentageCCCM[indexRatio] > 0){
        
        percentageText <- paste(percentageText, "The average value of Controls changed to Case is a **", round(totalTable$percentageCCCM[indexRatio], digits = 1), "%", " higher** than that of the Controls held as controls." , sep = "")
        
      } else {
        
        percentageText <- paste(percentageText, "The average value of Controls changed to Case is a **", abs(round(totalTable$percentageCCCM[indexRatio], digits = 1)), "%", " less** than that of the Controls held as controls.", sep = "")
        
      }
    }
    
    
    
  } else {

    predictor <- strsplit(rownames(totalTable)[indexRatio], " vs ")[[1]][1]
    
    allFlag = FALSE
    
    if (grepl("all =", predictor)) {
      allFlag = TRUE
    }
    
    if(!allFlag){
      predictorValue <- strsplit(predictor, " ")[[1]][2]
      predictorValue <- replace(predictorValue, is.na(predictorValue), "Empty")
      
      vsValue <- strsplit(rownames(totalTable)[indexRatio], " vs ")[[1]][2]
      vsValue <- replace(vsValue, is.na(vsValue), "Empty")
    } else {
      
      predictorValue <- strsplit(predictor, " ")[[1]][4]
      predictorValue <- replace(predictorValue, is.na(predictorValue), "Empty")
      
    }
    
    

    text <- paste(strsplit(predictor, " ")[[1]][1], "=", predictorValue, "(Categorical)", "<br>", sep = " ")
    
    if(allFlag){
      
      text <- paste(text, "Ratio:", round(totalTable$CCCM[indexRatio], digits = 3), sep = " ")
      
    } else {
      
      text <- paste(text, "Ratio: ", round(totalTable$CCCM[indexRatio], digits = 3), " -> (", totalTable$SumCCNumerator[indexRatio], "/", totalTable$SumCCDenominator[indexRatio], ") / (", totalTable$SumCMDenominator[indexRatio], "/", totalTable$SumCMNumerator[indexRatio], ")",  sep = "")
      
    }
    
    text <- paste(text, "Ratio (logaritmic scale):", sep = "<br>")
    text <- paste(text, round(totalTable$logCCCM[indexRatio], digits = 3),  sep = " ")

    text <- paste(text, "<br>", sep = "")
    text <- paste(text, "Number of Control -> Case patients in ", predictorValue, ":", sep = "")

    text <- paste(text, totalTable$SumCCNumerator[indexRatio], sep = " ")
    text <- paste(text, "<br>", sep = "")
    
    if(!allFlag){
      text <- paste(text, "Number of Control -> Case patients in ", vsValue, ":", sep = "")
  
      text <- paste(text, totalTable$SumCCDenominator[indexRatio], sep = " ")
      text <- paste(text, "<br>", sep = "")
    }
    
    text <- paste(text, "Number of Control -> Control patients in ", predictorValue, ":", sep = "")
    text <- paste(text, totalTable$SumCMNumerator[indexRatio], sep = " ")

    if(!allFlag){
      text <- paste(text, "<br>", sep = "")
      
      text <- paste(text, "Number of Control -> Control patients in ", vsValue, ":", sep = "")
  
      text <- paste(text, totalTable$SumCMDenominator[indexRatio], sep = " ")
    }
    
    percentageText <- paste("The variable **", rownames(totalTable)[indexRatio], "** is **categorical**.", sep = "")
      
    percentageText <- paste(percentageText, "\n", sep ="")
    
    percentageText <- paste(percentageText, "When the values **", predictorValue, " and ", vsValue, " are confronted**, a **", round(totalTable$CCCM[indexRatio], digits = 3), " ratio** is obtained.", sep = "")
    
    if(totalTable$CCCM[indexRatio] != 0){
      
      percentageText <- paste(percentageText, "This ratio means:", sep = " ")
      
      percentageText <- paste(percentageText, "\n", sep ="")
      
      if (totalTable$percentageCCCM[indexRatio] > 0){
        
        percentageText <- paste(percentageText, "That there are **", round(totalTable$percentageCCCM[indexRatio], digits = 1), "% more patients in ", predictorValue, "** that have been changed from Control to Case than in ", vsValue, "." , sep = "")
        
      } else {
        
        percentageText <- paste(percentageText, "That there are **", abs(round(totalTable$percentageCCCM[indexRatio], digits = 1)), "% less patients in ", predictorValue, "** that have been changed from Control to Case than in ", vsValue, "." , sep = "")
        
      }
    
    }

  }

  totalTable$textCCCM[indexRatio] <- text
  totalTable$percentageTextCCCM[indexRatio] <- percentageText
  
  if(totalTable$ECEM[indexRatio] == 0){

    totalTable$ColorsECEM[indexRatio] <- "black"

  } else {

    if(abs(totalTable$logECEM[indexRatio]) < 0.5){
      totalTable$ColorsECEM[indexRatio] <- "gold"
    } else if (abs(totalTable$logECEM[indexRatio]) > 0.5 && abs(totalTable$logECEM[indexRatio]) < 1.5){
      totalTable$ColorsECEM[indexRatio] <- "orange"
    } else {
      totalTable$ColorsECEM[indexRatio] <- "red"
    }

  }

  text <- ""
  percentageText <- ""

  if(rownames(totalTable)[indexRatio] %in% numericVariables){

    text <- strsplit(rownames(totalTable)[indexRatio], " vs ")[[1]][1]

    text <- paste(text, "(Numeric)", "<br>", sep = " ")
      
    text <- paste(text, "Ratio: ", round(totalTable$ECEM[indexRatio], digits = 3), " -> (", round(numericTable[rownames(totalTable)[indexRatio], "MeanEC"], digits = 3), "/", round(numericTable[rownames(totalTable)[indexRatio], "MeanEM"], digits = 3), ")",  sep = "")
  
    text <- paste(text, "Ratio (logaritmic scale):", sep = "<br>")
    text <- paste(text, round(totalTable$logECEM[indexRatio], digits = 3),  sep = " ")
    
    text <- paste(text, "Number of Case -> Control patients:", sep = "<br>")

    text <- paste(text, numericTable[rownames(totalTable)[indexRatio], "NumPatientsEC"], sep = " ")


    text <- paste(text, "Mean value for the Case -> Control patients:", sep = "<br>")

    text <- paste(text, round(numericTable[rownames(totalTable)[indexRatio], "MeanEC"], digits = 3), sep = " ")


    text <- paste(text, "Number of Case -> Case patients:", sep = "<br>")

    text <- paste(text, numericTable[rownames(totalTable)[indexRatio], "NumPatientsEM"], sep = " ")


    text <- paste(text, "Mean value for the Case -> Case patients:", sep = "<br>")

    text <- paste(text, round(numericTable[rownames(totalTable)[indexRatio], "MeanEM"], digits = 3), sep = " ")
    
    
    percentageText <- paste("The variable **", rownames(totalTable)[indexRatio], "** is **numeric**.", sep = "")
      
    percentageText <- paste(percentageText, "\n", sep ="")
    

    percentageText <- paste(percentageText, "A **", round(totalTable$ECEM[indexRatio], digits = 3), " ratio** is obtained.", sep = "")
    
    if(totalTable$ECEM[indexRatio] != 0){
      
      percentageText <- paste(percentageText, "This ratio means:", sep = " ")
      
      percentageText <- paste(percentageText, "\n", sep ="")
      
      if (totalTable$percentageECEM[indexRatio] > 0){
        
        percentageText <- paste(percentageText, "The average value of Controls changed to Case is a **", round(totalTable$percentageECEM[indexRatio], digits = 1), "%", " higher** than that of the Controls held as controls.", sep = "")
        
      } else {
        
        percentageText <- paste(percentageText, "The average value of Controls changed to Case is a **", abs(round(totalTable$percentageECEM[indexRatio], digits = 1)), "%", " less** than that of the Controls held as controls.", sep = "")
        
      }
    }
    
  } else {

    predictor <- strsplit(rownames(totalTable)[indexRatio], " vs ")[[1]][1]
    
    allFlag = FALSE
    
    if (grepl("all =", predictor)) {
      allFlag = TRUE
    }
    
    if(!allFlag){
      predictorValue <- strsplit(predictor, " ")[[1]][2]
      predictorValue <- replace(predictorValue, is.na(predictorValue), "Empty")
      
      vsValue <- strsplit(rownames(totalTable)[indexRatio], " vs ")[[1]][2]
      vsValue <- replace(vsValue, is.na(vsValue), "Empty")
    } else {
      predictorValue <- strsplit(predictor, " ")[[1]][4]
      predictorValue <- replace(predictorValue, is.na(predictorValue), "Empty")
    }
    
    
    text <- paste(strsplit(predictor, " ")[[1]][1], "=", predictorValue, "(Categorical)","<br>", sep = " ")
    
    if(allFlag){
      
      text <- paste(text, "Ratio:", round(totalTable$ECEM[indexRatio], digits = 3), sep = " ")
      
    } else {
      
      text <- paste(text, "Ratio: ", round(totalTable$ECEM[indexRatio], digits = 3), " -> (", totalTable$SumECNumerator[indexRatio], "/", totalTable$SumECDenominator[indexRatio], ") / (", totalTable$SumEMNumerator[indexRatio], "/", totalTable$SumEMDenominator[indexRatio], ")",  sep = "")
      
    }
    
    text <- paste(text, "Ratio (logaritmic scale):", sep = "<br>")
    text <- paste(text, round(totalTable$logECEM[indexRatio], digits = 3),  sep = " ")

    text <- paste(text, "<br>", sep = "")
    text <- paste(text, "Number of Case -> Control patients in ", predictorValue, ":", sep = "")

    text <- paste(text, totalTable$SumECNumerator[indexRatio], sep = " ")
    text <- paste(text, "<br>", sep = "")
    
    if(!allFlag){
      text <- paste(text, "Number of Case -> Control patients in ", vsValue, ":", sep = "")
  
      text <- paste(text, totalTable$SumECDenominator[indexRatio], sep = " ")
  
      text <- paste(text, "<br>", sep = "")
    }

    text <- paste(text, "Number of Case -> Case patients in ", predictorValue, ":", sep = "")

    text <- paste(text, totalTable$SumEMNumerator[indexRatio], sep = " ")

    
    
    if(!allFlag){
      text <- paste(text, "<br>", sep = "")
      
      text <- paste(text, "Number of Case -> Case patients in ", vsValue, ":", sep = "")
  
      text <- paste(text, totalTable$SumEMDenominator[indexRatio], sep = " ")
    }
     
    percentageText <- paste("The variable **", rownames(totalTable)[indexRatio], "** is **categorical**.", sep = "")
      
    percentageText <- paste(percentageText, "\n", sep ="")
    
    percentageText <- paste(percentageText, "When the values **", predictorValue, " and ", vsValue, " are confronted**, a **", round(totalTable$ECEM[indexRatio], digits = 3), " ratio** is obtained.", sep = "")
    
    if(totalTable$ECEM[indexRatio] != 0){
      
      percentageText <- paste(percentageText, "This ratio means:", sep = " ")
      
      percentageText <- paste(percentageText, "\n", sep ="")
      
      if (totalTable$percentageECEM[indexRatio] > 0){
        
        percentageText <- paste(percentageText, "That there are **", round(totalTable$percentageECEM[indexRatio], digits = 1), "% more patients in ", predictorValue, "** that have been changed from Control to Case than in ", vsValue, "." , sep = "")
        
      } else {
        
        percentageText <- paste(percentageText, "That there are **", abs(round(totalTable$percentageECEM[indexRatio], digits = 1)), "% less patients in ", predictorValue, "** that have been changed from Control to Case than in ", vsValue, "." , sep = "")
        
      }
    }
  }

  totalTable$textECEM[indexRatio] <- text
  totalTable$percentageTextECEM[indexRatio] <- percentageText
  
  if(totalTable$CE[indexRatio] == 0){

    totalTable$ColorsCE[indexRatio] <- "black"

  } else {

    if(abs(totalTable$logCE[indexRatio]) < 0.5){
      totalTable$ColorsCE[indexRatio] <- "gold"
    } else if (abs(totalTable$logCE[indexRatio]) > 0.5 && abs(totalTable$logCE[indexRatio]) < 1.5){
      totalTable$ColorsCE[indexRatio] <- "orange"
    } else {
      totalTable$ColorsCE[indexRatio] <- "red"
    }

  }

  text <- ""
  percentageText <- ""
  
  if(rownames(totalTable)[indexRatio] %in% numericVariables){

    text <- strsplit(rownames(totalTable)[indexRatio], " vs ")[[1]][1]

    text <- paste(text, "(Numeric)", "<br>", sep = " ")
  
    text <- paste(text, "Ratio: ", round(totalTable$CE[indexRatio], digits = 3), " -> (", round(numericTable[rownames(totalTable)[indexRatio], "MeanC"], digits = 3), "/", round(numericTable[rownames(totalTable)[indexRatio], "MeanE"], digits = 3), ")",  sep = "")

    text <- paste(text, "Ratio (logaritmic scale):", sep = "<br>")
    text <- paste(text, round(totalTable$logCE[indexRatio], digits = 3),  sep = " ")
    
    text <- paste(text, "Number of Control patients (before genetic):", sep = "<br>")

    text <- paste(text, numericTable[rownames(totalTable)[indexRatio], "NumPatientsC"], sep = " ")


    text <- paste(text, "Mean value for the Control patients (before genetic):", sep = "<br>")

    text <- paste(text, round(numericTable[rownames(totalTable)[indexRatio], "MeanC"], digits = 3), sep = " ")


    text <- paste(text, "Number of Case patients (before genetic):", sep = "<br>")

    text <- paste(text, numericTable[rownames(totalTable)[indexRatio], "NumPatientsE"], sep = " ")


    text <- paste(text, "Mean value for the Case patients (before genetic):", sep = "<br>")

    text <- paste(text, round(numericTable[rownames(totalTable)[indexRatio], "MeanE"], digits = 3), sep = " ")
    
    
    percentageText <- paste("The variable **", rownames(totalTable)[indexRatio], "** is **numeric**.", sep = "")
      
    percentageText <- paste(percentageText, "\n", sep ="")
    
    percentageText <- paste(percentageText, "A **", round(totalTable$CE[indexRatio], digits = 3), " ratio** is obtained.", sep = "")
    
    
    if(totalTable$CE[indexRatio] != 0){

      percentageText <- paste(percentageText, "This ratio means:", sep = " ")
      
      percentageText <- paste(percentageText, "\n", sep ="")
      
      if (totalTable$percentageCE[indexRatio] > 0){
        
        percentageText <- paste(percentageText, "The average value of the Control patients before the detection is a **", round(totalTable$percentageCE[indexRatio], digits = 1), "%", " higher** than that of the Case patients.", sep = "")
        
      } else {
        
        percentageText <- paste(percentageText, "The average value of the Control patients before the detection is a **", abs(round(totalTable$percentageCE[indexRatio], digits = 1)), "%", " less** than that of the Case patients.", sep = "")
        
      }
    }
    
  } else {

    predictor <- strsplit(rownames(totalTable)[indexRatio], " vs ")[[1]][1]
    
    allFlag = FALSE
    
    if (grepl("all =", predictor)) {
      allFlag = TRUE
    }
    
    if(!allFlag){
      predictorValue <- strsplit(predictor, " ")[[1]][2]
      predictorValue <- replace(predictorValue, is.na(predictorValue), "Empty")
      
      vsValue <- strsplit(rownames(totalTable)[indexRatio], " vs ")[[1]][2]
      vsValue <- replace(vsValue, is.na(vsValue), "Empty")
    } else {
      
      predictorValue <- strsplit(predictor, " ")[[1]][4]
      predictorValue <- replace(predictorValue, is.na(predictorValue), "Empty")
    }
    
    
    
    text <- paste(strsplit(predictor, " ")[[1]][1], "=", predictorValue, "(Categorical)", "<br>", sep = " ")
    
    if(allFlag){
      
      text <- paste(text, "Ratio:", round(totalTable$CE[indexRatio], digits = 3), sep = " ")
      
    } else {
      
      text <- paste(text, "Ratio: ", round(totalTable$CE[indexRatio], digits = 3), " -> (", totalTable$SumCNumerator[indexRatio], "/", totalTable$SumCDenominator[indexRatio], ") / (", totalTable$SumENumerator[indexRatio], "/", totalTable$SumEDenominator[indexRatio], ")",  sep = "")
      
    }
    
    text <- paste(text, "Ratio (logaritmic scale):", sep = "<br>")
    text <- paste(text, round(totalTable$logCE[indexRatio], digits = 3),  sep = " ")
    
    text <- paste(text, "<br>", sep = "")
    
    text <- paste(text, "Number of Control patients (before genetic) in ", predictorValue, ":", sep = "")

    text <- paste(text, totalTable$SumCNumerator[indexRatio], sep = " ")

    text <- paste(text, "<br>", sep = "")
    
    if(!allFlag){
      text <- paste(text, "Number of Control patients (before genetic) in ", vsValue, ":", sep = "")
  
      text <- paste(text, totalTable$SumCDenominator[indexRatio], sep = " ")
  
      text <- paste(text, "<br>", sep = "")
    }

    text <- paste(text, "Number of Case patients (before genetic) in ", predictorValue, ":", sep = "")

    text <- paste(text, totalTable$SumENumerator[indexRatio], sep = " ")
    
    if(!allFlag){
      text <- paste(text, "<br>", sep = "")
  
      text <- paste(text, "Number of Case patients (before genetic) in ", vsValue, ":", sep = "")
  
      text <- paste(text, totalTable$SumEDenominator[indexRatio], sep = " ")
    }
    
    
    percentageText <- paste("The variable **", rownames(totalTable)[indexRatio], "** is **categorical**.", sep = "")
      
    percentageText <- paste(percentageText, "\n", sep ="")
    
    percentageText <- paste(percentageText, "When the values **", predictorValue, " and ", vsValue, " are confronted**, a **", round(totalTable$CE[indexRatio], digits = 3), " ratio** is obtained.", sep = "")
    
    if(totalTable$CE[indexRatio] != 0){

      percentageText <- paste(percentageText, "This ratio means:", sep = " ")
      
      percentageText <- paste(percentageText, "\n", sep ="")
      
      if (totalTable$percentageCE[indexRatio] > 0){
        
        percentageText <- paste(percentageText, "That there are **", round(totalTable$percentageCE[indexRatio], digits = 1), "% more Control patients in ", predictorValue, " that Control patients in ", vsValue, "**." , sep = "")
        
      } else {
        
        percentageText <- paste(percentageText, "That there are **", abs(round(totalTable$percentageCE[indexRatio], digits = 1)), "% less Control patients in ", predictorValue, " that Control patients in ", vsValue, "**." , sep = "")
        
      }
    }
  }

  totalTable$textCE[indexRatio] <- text
  totalTable$percentageTextCE[indexRatio] <- percentageText
}
```

### 9.1 Ratios between Control and Case patients before detection

For **numerical** attributes, on the **right** we see those that show **higher mean values for controls than for cases**. Analogously on the left. 

For **categorical** attributes, we see a blob for each possible category versus the rest. If the attribute is binary, only a blob is required. We see there the **ratio of members of one category and the rest**, for controls, **divided by the ratio of members of that category and the rest**, for cases.

```{r, include = FALSE}
plotCE <- plot_ly(totalTable,x=~logCE,y=~rownames(totalTable),type="scatter", mode = "markers", marker = list(size = 10, color = totalTable$ColorsCE), height = (nrow(totalTable) * 850) / 36, width = 1000, hoverinfo = "text", text =~textCE)

title <- "Ratios between Control and Case patients before detection (genetic algorithm)"

plotCE <- plotCE %>% layout(title = title,
                        xaxis = list(title = "Ratio in logaritmic (base 2) scale <br>", tickvals = seq(-ceiling(-min(totalTable$logCE)), ceiling(max(totalTable$logCE)), by = sepCE)),
                        yaxis = list(title = "Predictor"))
```

```{r}
plotCE
```

`r paste(totalTable$percentageTextCE, collapse = "\n\n")`

### 9.2 Evolution of the set of Control patients after detection

For **numerical** attributes, on the **right** we see those that show **higher mean values for controls that became cases in comparison to controls that remained as controls**. Analogously on the left 

For **categorical** attributes, we see a blob for each possible category versus the rest. If the attribute is binary, only a blob is required. We see there the **ratio of members of one category versus the rest**, for all individuals that changed from control to cases, **divided by the ratio of members of that category versus the rest**, that remained as controls.

```{r, include = FALSE}
plotCCCM <- plot_ly(totalTable,x=~logCCCM,y=~rownames(totalTable), type="scatter", mode = "markers", marker = list(size = 10, color = totalTable$ColorsCCCM ), height = (nrow(totalTable) * 850) / 36, width = 1000, hoverinfo = "text", text =~textCCCM)

title <- "Evolution of the set of Control patients after detection (genetic algorithm)"
plotCCCM <- plotCCCM %>% layout(title = title,
                        xaxis = list(title = "Ratio in logaritmic (base 2) scale <br>", tickvals = seq(-ceiling(-min(totalTable$logCCCM)), ceiling(max(totalTable$logCCCM)), by = sepCCCM)),
                        yaxis = list(title = "Predictor"))
```

```{r}
plotCCCM
```

`r paste(totalTable$percentageTextCCCM, collapse = "\n\n")`

### 9.3 Evolution of the set of Case patients after detection

For **numerical** attributes, on the **right** we see those that show **higher mean values for cases that became controls in comparison to cases that remained as cases**. Analogously on the left 

For **categorical** attributes, we see a blob for each possible category versus the rest. If the attribute is binary, only a blob is required. We see there the **ratio of members of one category versus the rest**, for all individuals that changed from cases to controls, **divided by the ratio of members of that category versus the rest**, that remained as cases.

```{r, include = FALSE}
plotECEM <- plot_ly(totalTable,x=~logECEM,y=~rownames(totalTable),type="scatter", mode = "markers", marker = list(size = 10, color = totalTable$ColorsECEM ), height = (nrow(totalTable) * 850) / 36, width = 1000, hoverinfo = "text", text =~textECEM)

title <- "Evolution of the set of Case patients after detection (genetic algorithm)"

plotECEM <- plotECEM %>% layout(title = title,
                        xaxis = list(title = "Ratio in logaritmic (base 2) scale <br>", tickvals = seq(-ceiling(-min(totalTable$logECEM)), ceiling(max(totalTable$logECEM)), by = sepECEM)),
                        yaxis = list(title = "Predictor"))
```

```{r}
plotECEM
```

`r paste(totalTable$percentageTextECEM, collapse = "\n\n")`
